---
title: "Assigment - Naive Bayes DIY"
author:
  - name author here - Author
  - name reviewer here - Reviewer
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
  html_notebook:
    toc: yes
    toc_depth: 2
---

```{r results = "hide"}
library(tidyverse)
library(tm)
library(caret)
library(wordcloud)
library(e1071)
```

---

Choose a suitable dataset from [this](https://github.com/HAN-M3DM-Data-Mining/assignments/tree/master/datasets) folder and train your own Naive Bayes model. Follow all the steps from the CRISP-DM model.

**Chosen dataset**: Fake News


## Business Understanding
text and code here

## Data Understanding

The dataset was retrived from Kaggle, which is a community-maintained repository for public dataset. The data is comprised of news articles (the instances) along with labels for those articles (trustworthy or not) and a few other variables.

First we will look at the first 10 rows of the dataset to get a better understanding of the data.

```{r}
fakeNews_url = 'https://raw.githubusercontent.com/HAN-M3DM-Data-Mining/assignments/master/datasets/NB-fakenews.csv'
rawDF = read_csv(fakeNews_url)
head(rawDF) # returns an error for some reason, works fine in the console
```

We can see that the dataset has five columns. The labels are market with 1 (unreliable) and 0 (reliable). The other variables are the ID of the article (numeric), title (character), author (character) and the content of the aricle in the text column (character).

In the summary of the dataframe we can see that there are 20800 observations.

```{r}
summary(rawDF)
```
Let us also examine the most commonly occuring words in reliable and unreliable articles through a wordcloud.

```{r}
unreliable = rawDF %>% filter(label == 1)
reliable = rawDF %>% filter(label == 0)

wordcloud(unreliable$text, max.words = 20, scale = c(4, 0.8), colors= c("indianred1","indianred2","indianred3","indianred"))
wordcloud(reliable$text, max.words = 20, scale = c(4, 0.8), colors= c("lightsteelblue1","lightsteelblue2","lightsteelblue3","lightsteelblue"))
```



## Data Preparation
text and code here

## Modeling
text and code here

## Evaluation and Deployment
text and code here

reviewer adds suggestions for improving the model